Rapid, Unsupervised Generation of Novel Percussive Sounds With A Tractable Virtual Synthesizer

### 2020-01-01 NIME_PAPER/paper.tex
Updates from Overleaf

\begin{abstract}
With the generation of novel, high quality, one shot percussive sounds as our end goal, we implemented a set of modular software components utilized to form a pipeline for guided generation of sounds. The most crucial of these components are a "virtual ear" capable of evaluating a sounds proximity to different drum types and an efficient, tractable "virtual synthesizer" with a rich set of parameters and DSP methods capable of generating a wide range of sounds.\\ 
We aimed to design our software such that we can utilize our components in various experimental ways; hence, in order to established an adequate level of confidence in our components, we discuss the implementation, limitations, and evaluations of the \textit{virtual synthesizer} and \textit{virtual ear}. Next, we present our findings and assessments in regards to the three different methodologies utilized for the generative process of percussive sounds:  \begin {enumerate*} [label=(\roman*)]
\item random generation \item multivariate distribution sampling and \item evolutionary search.\\
\end {enumerate*} 
Additionally, as we observed a general lack of copyright-free percussive sample-packs as well as a lack of unified approach for obtaining such a data-set by fellow researchers \cite{aouameur2019neural,ramires2019timbfeat}, we aim to ameliorate the issue by presenting a copyright free sample-pack of categorized percussion sounds. Furthermore, we have included in our project source code \footnote{\url{https://github.com/imilas/Synths_Stacks_Search}} a set of scripts to further expand the sample pack via royalty free sources. 
\keywords{drum, percussion, synthesis, MIR, ML, genetic, search}
\end{abstract}
% See template for CCS classification. 
\ccsdesc[500]{Applied computing~Sound and music computing}
\ccsdesc[500]{Information systems~Music retrieval}
\printccsdesc

\section{Introduction}
\subsection{Terminology}
Will define terms such as VSTs, audio effects, Synthesizers, paramters and presets here. I throw these terms around a lot and I think some might be confused by their meaning.
\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly refered to as pads or textures), percussive or melodic loops etc. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore VST plug-ins can emulate complex audio synthesizers and effects, which some producers may find daunting or time consuming to work with from scratch. In many cases, VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. Users may modify any loaded preset until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers that a producer might find  interesting. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for promising avenues towards our motivational goal.\\

@@

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used LSTM and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source, making the eventual replication certain, even with random search. Feature matching also appears to not be concerned with the envelope of the sounds, but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio. As stated before, our work presents a rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer.
\section{Implementation}
\label{impl}
\subsection{Design}

>>>>

\begin{abstract}
With the generation of novel, high quality, one shot percussive sounds as our end goal, we implemented a set of modular software components utilized to form a pipeline for guided generation of sounds. The most crucial of these components are a "virtual ear" capable of evaluating a sound's proximity to different drum types and an efficient, tractable "virtual synthesizer" with a rich set of parameters capable of generating a wide range of sounds.\\ 
We aimed to design our software such that we can utilize our components in various experimental ways; hence, in order to established an adequate level of confidence in our components, we discuss the implementation, limitations, and evaluations of the \textit{virtual synthesizer} and \textit{virtual ear}. Next, we present our findings and assessments in regards to the three different methodologies utilized for the generative process of percussive sounds:  \begin {enumerate*} [label=(\roman*)]
\item random generation \item multivariate distribution sampling and \item evolutionary search.\\
\end {enumerate*} 
Additionally, as we observed a general lack of copyright-free percussive sample-packs as well as a lack of unified approach for obtaining such a data-set by fellow researchers \cite{aouameur2019neural,ramires2019timbfeat}, we aim to ameliorate the issue by presenting a copyright free sample-pack of categorized percussion sounds. Furthermore, we have included in our project source code \footnote{\url{https://github.com/imilas/Synths_Stacks_Search}} a set of scripts to further expand the sample pack via royalty free sources. 
\keywords{drum, percussion, synthesis, MIR, ML, genetic, search}
\end{abstract}
% See template for CCS classification. 
\ccsdesc[500]{Applied computing~Sound and music computing}
\ccsdesc[500]{Information systems~Music retrieval}
\printccsdesc

\section{Introduction}
\subsection{Terminology}
Will define terms such as VSTs, audio effects, Synthesizers, paramters and presets here. I throw these terms around a lot and I think some might be confused by their meaning.
\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops(percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects, which some producers may find daunting or time consuming to work with from scratch. In many cases, VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. Producers may further modify these presets until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers that a producer might find  interesting. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for promising avenues towards our motivational goal.\\

@@

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used LSTM and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain, even with random search. Feature matching also appears to not be concerned with the envelope of the sounds, but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio. As stated before, our work presents a rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer.
\section{Implementation}
\label{impl}
\subsection{Design}
